---
title: "Lab 5 Time Series & Spatial Analysis"
author: "Jonathan Hart"
date: "2/6/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

## Attach packages
```{r}
library(tidyverse)
library(janitor)
library(lubridate)
library(here)
library(paletteer)
library(tsibble)
library(fable)
library(fabletools)
library(feasts)
library(forecast)
library(sf)
library(tmap)
library(mapview)
```

## Monthly US energy consumption (renewables)
```{r}
us_renew <- read_csv(here("data", "renewables_cons_prod.csv")) %>% 
  clean_names()
```

- Make desciption all lowercase 
- only keep observations where "description" variable contains "consumption"
- remove any observations where "description" variable contains "total"

```{r}
renew_clean <- us_renew %>% 
  mutate(description = str_to_lower(description)) %>% 
  filter(str_detect(description, pattern = "consumption")) %>% 
  filter(!str_detect(description, pattern = "total")) 
```

### Convert 'yyyymm' column to a date 
```{r}
renew_date <- renew_clean %>% 
  mutate(yr_mo_day = lubridate::parse_date_time(yyyymm, "ym")) %>% 
  mutate(month_sep = yearmonth(yr_mo_day)) %>% 
  mutate(value = as.numeric(value)) %>% 
  drop_na(month_sep, value) 

# Make a version where I hace the month & year in separate columns
renew_parsed <- renew_date %>% 
  mutate(month = month(yr_mo_day, label = TRUE)) %>% 
  mutate(year = year(yr_mo_day))
```

### Look at it:

```{r}
renew_gg <- ggplot(data = renew_date, aes(x = month_sep, 
                                          y = value,
                                          group = description)) +
  geom_line(aes(color = description))

renew_gg
```

Updating colors with palatter palettes:
```{r}
renew_gg +
  scale_color_paletteer_d("calecopal::oak")
```

### Coerce renew_parsed to a tsibble

```{r}
renew_ts <- as_tsibble(renew_parsed, key = description, index = month_sep)
```

Let's look at our ts data in a couple different ways:
```{r}
renew_ts %>% autoplot(value)
renew_ts %>% gg_subseries(value)
renew_ts %>%  gg_season(value)

# ggplot(data = renew_parsed, aes(x = month, y = value, group = year)) +
#  geom_line(aes(color = year)) +
#  facet_wrap(~description,
#             ncol = 1,
#             scales = "free",
#             strip.position = "right")
```

### Just look at the hydroelectric energy consumption
```{r}
hydro_ts <- renew_ts %>% 
  filter(description == "hydroelectric power consumption")


hydro_ts %>% autoplot(value)
hydro_ts %>% gg_subseries(value)
hydro_ts %>% gg_season(value)
```
### Whar if I want quarterly average for hydro?
```{r}
hydro_annual <- hydro_ts %>% 
  index_by(annual = ~year(.)) %>% 
  summarize(
    avg_consumption = mean(value)
  )

ggplot(data = hydro_annual, aes(x = annual, y = avg_consumption)) +
  geom_line(color = "darkorange") +
  geom_smooth(color = "purple",
              size = 0.2,
              linetype = "dashed",
              fill = "purple",
              alpha = 0.2) +
  theme_minimal()
```

### Decompose the hydro consumption ts data

First, let's check the decomposition (STL):
```{r}
# Find STL decomposition
dcmp <- hydro_ts %>%
  model(STL(value ~ season(window = 5)))

# View the components
# components(dcmp)

# Visualize the decomposed components
components(dcmp) %>% autoplot() +
  theme_minimal()

# Let's check out the residuals:
hist(components(dcmp)$remainder)

```


### Explore the ACF
```{r}

hydro_ts %>% 
  ACF(value) %>% 
  autoplot()

```
We see peaks at 12 months: annual-difference similarities in consumption. 

### Forecast future hydro power consumption 

```{r}
hydro_model <- hydro_ts %>%
  model(
    arima = ARIMA(value)
  ) %>%
  fabletools::forecast(h = "2 years")

hydro_model %>% autoplot(filter(hydro_ts, year(month_sep) > 2010), level = NULL)

```
### Make a world map!

```{r}
world <- read_sf(dsn = here("data", "TM_WORLD_BORDERS_SIMPL-0.3-1"),
                 layer = "TM_WORLD_BORDERS_SIMPL-0.3")

mapview(world)
```

